---
title: "dataRS_at_a_glance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dataRS_in_a_glimpse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 10, 
  fig.height = 10,
  comment = "#>"
)
```

#dataRS package at a glance

##1) Install and load
```{r setup, warning = FALSE, message = FALSE, results = FALSE, echo = TRUE}
#Installation from github
library(devtools)
#devtools::install_github("fernandoilcruz/dataRS")

#Load the package
library(dataRS)
```

##2) Set API key
```{r warning = FALSE, message = FALSE, results = FALSE, echo = TRUE}
#You can hardcode your API key (not recommended)
#dataRS::set_api_key("your_api_key_here")

#Or save it in your .Renviron file as a global variable (more appropriate)
my_key <- Sys.getenv("MY_API_KEY")
dataRS::set_api_key(key = my_key)
```

##3) Select variables by ID
For this example, we will collect all IDs related to the population estimates by RIPSA.
```{r warning = FALSE, message = FALSE, results = FALSE, echo = TRUE}
my_vars <- 
  dataRS::vars() |>
  dplyr::filter(stringr::str_detect(var_name,"RIPSA")) |>
  dplyr::select(var_id) |>
  dplyr::pull()
```

##4) Select geographic region by ID
For this example, we will select Coredes Hortênsias and Metropolitano-Delta do Jacuí.
```{r warning = FALSE, message = FALSE, results = FALSE, echo = TRUE}
my_geo <- 
  dataRS::geoagreg(ag = "corede") |>
  dplyr::filter(stringr::str_detect(geo_name,"Hortênsias") | stringr::str_detect(geo_name,"Metropolitano")) |>
  dplyr::select(geo_id) |>
  dplyr::pull()
```

#5) Check if data is available for your selections (optional)
```{r warning = FALSE, message = FALSE, results = FALSE, echo = TRUE}
my_vars |>
  purrr::set_names() |>
  purrr::map(
    function(x){
      dataRS::available(x, ag = "corede")
      }
    )
```

##6) Download data
```{r warning = FALSE, message = FALSE, results = FALSE, echo = TRUE}
my_data <- 
  dataRS::getdata(var_id = my_vars, ag = "corede", geo_id = my_geo, period = "all", add_labels = TRUE)
```


##EXTRA
The dataRS package also provides additional regionalization schemes that can be linked to municipalities.

```{r warning = FALSE, message = FALSE, results = FALSE, echo = TRUE}
data("geo_complement")
geo_complement
```

